---
title: "R Notebook"
output: html_notebook
---

```{r}
library(RPostgreSQL)
```

** 1 Preparation **

Connect to local database

```{r}
home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)
```

```{r}
db_connect <- dbConnect(drv = PostgreSQL(), user = postgresql_username, password = postgresql_password, dbname = "omni_employees", host = "localhost")
```

Remove username and password from list of values for privacy

```{r}
rm(postgresql_username, postgresql_password)
```

List tables in omni_employees db

```{r}
dbListTables(conn = db_connect)
```

List fields in the employees table

```{r}
dbListFields(conn = db_connect, "employees")
```

Query employees table to view data

```{r}
dbGetQuery(conn = db_connect, statement = "SELECT * FROM employees")
```
** 2 MVP **

1. Find all the employees who work in the ‘Human Resources’ department

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE department = 'Human Resources'
```
2. Get the first name, last name, and country of the employees who work in the ‘Legal’ department

```{sql connection=db_connect}
SELECT
  first_name,
  last_name,
  country
FROM employees
WHERE department = 'Legal'
```
3. Count the number of employees based in Portugal

```{sql connection=db_connect}
SELECT
  COUNT(id) AS num_employees
FROM employees
WHERE country = 'Portugal'
```

4. Count the number of employees based in either Portugal or Spain

```{sql connection=db_connect}
SELECT
  COUNT(id) AS num_employees
FROM employees
WHERE country IN ('Portugal', 'Spain')
```

5. Count the number of pay_details records lacking a local_account_no

  5.1 First query pay_details table to view data

```{r}
dbGetQuery(conn = db_connect, statement = "SELECT * FROM pay_details")
```
  5.1 SQL to answer question

```{sql connection=db_connect}
SELECT
  COUNT(id) AS num_pay_details
FROM pay_details
WHERE local_account_no IS NULL
```

6. Are there any pay_details record lacking both a local_account_no and iban number?

```{sql connection=db_connect}
SELECT *
FROM pay_details
WHERE local_account_no IS NULL AND iban IS NULL
```

7. Get a list of employees first names and last names ordered alphabetically by their last name (put any nulls last)

```{sql connection=db_connect}
SELECT
  first_name,
  last_name 
FROM employees
ORDER BY last_name ASC NULLS LAST
```
  There are 5 last_name with lower case first character, which are incorrectly ordered last

  Use LOWER to make last_name lower case to ORDER BY

```{sql connection=db_connect}
SELECT
  first_name,
  last_name
FROM employees
ORDER BY LOWER(last_name) ASC NULLS LAST
```

8. Get a list of employees first names, last names and countries, ordered alphabetically first by their country and second by their last name (put any nulls last)

```{sql connection=db_connect}
SELECT
  first_name,
  last_name,
  country
FROM employees
ORDER BY country, LOWER(last_name) ASC NULLS LAST
```

9. Find the top ten highest paid employees in the corporation

```{sql connection=db_connect}
SELECT *
FROM employees
ORDER BY salary DESC NULLS LAST
LIMIT 10
```
10. Find the first name, last name and salary of the lowest paid of the employees in Hungary

```{sql connection=db_connect}
SELECT
  first_name,
  last_name,
  salary
FROM employees
WHERE country = 'Hungary'
ORDER BY salary ASC NULLS LAST
LIMIT 1
```

11. How many employees have a first name beginning with ‘F’

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE first_name LIKE 'F%'
```
12. Find all the details of any employees with a ‘yahoo’ e-mail address

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE email LIKE '%yahoo%'
```

13. Count the number of pension enrolled employees not based in either France or Germany

```{sql connection=db_connect}
SELECT
  COUNT(id) AS num_employees
FROM employees
WHERE pension_enrol = TRUE AND country NOT IN ('France', 'Germany')
```

14. Breakdown the numbers of employees enrolled, not enrolled and whose enrollment status is unknown in the corporation pension scheme

```{sql connection=db_connect}
SELECT
  pension_enrol,
  COUNT(id) AS num_employees
FROM employees
GROUP BY pension_enrol
```

15. What is the maximum salary among employees in the Engineering department who work 1.0 full-time equivalent hours?

```{sql connection=db_connect}
SELECT salary
FROM employees
WHERE department = 'Engineering' AND fte_hours = 1.0
ORDER BY salary DESC
LIMIT 1
```

16. Obtain a count by department of the employees who started work with the corporation in 2003

```{sql connection=db_connect}
SELECT
  department,
  COUNT(id) AS num_employees
FROM employees
WHERE EXTRACT(YEAR FROM start_date) = 2003
GROUP BY department
```

Disconnect from local database

```{r}
dbDisconnect(db_connect)
```