---
title: "R Notebook"
output: html_notebook
---

** Question 1 **

Let’s start by loading the tidyverse library and reading the data in. You learned earlier this week how to get a feel for your data when you first load it in. Find out the dimensions, variable names, and print the first 10 rows.

Load libraries

```{r}
library(tidyverse)
library(janitor)
```

Read data

```{r}
face_descriptions <- read_csv("data/face_descriptions.csv")
```

Review data

```{r}
head(face_descriptions, 10)
```
```{r}
dim(face_descriptions)
```

** Question 2 **

Do these variable names look tidy to you? What format is your data in (long or wide)?

The column names are not tidy. They are not descriptive of the column values. The data is wide format which is a disadvantage over long (vertical) due to scrolling etc.

** Question 3 **

Being faced with such complex data can be daunting, we may feel overwhelmed and ask ourselves:

How can I organise this data?
How can I make this data meaningful?
How can I make this data tidy?
The first problem you can always tackle is to get your data in an appropriate format. What format do you need?

Once you have figured that out, reshape the data into the appropriate format.

Hint Use the pivot_longer() function to gather the data into long format. You’ll need to put all columns beginning with t into one column.

```{r}
face_descriptions_tidy <- face_descriptions %>%
  pivot_longer(cols = c("t1":"t50"),
               names_to = "t-value",
               values_to = "description") %>%
  select(-'t-value')

face_descriptions_tidy
```

** Question 4 **

But wait! Some people have given lengthy descriptions of the faces, some have only given one word. This can be problematic for analysis, if you eventually want to summarise key descriptions.

Some people have split the description of the faces using a slash. Use the separate function to split apart the descriptions column so that if people have given more than one answer (i.e. if the data is separated by a /) it is moved into a new row.

First, you’ll need to decide a cut off for how many responses you want to accept from people. Is one enough? Two? Three? Once you’ve decided how many descriptions you want to code, you’ll have to separate your description columns into that many columns.

```{r}
unique(face_descriptions_tidy$description)
```

```{r}
face_descriptions_clean <- face_descriptions_tidy %>%
  separate(col = description,
           into = c("description_1", "description_2", "description_3"),
           sep = "/")

face_descriptions_clean
```

```{r}
unique(face_descriptions_clean$description_3)
```

** Question 5 **

We’ve now split the data, and have three description variables. But is this data ACTUALLY tidy? Isn’t one of the rules that we need to have only one column per variable? And now we have more than one description variables…

What do we need to do here so our description variables follow the rules of tidy data?

Hint Use the pivot_longer() function to create a new description column, which will identify which description number it is (1,2,3, etc), and to create one single variable called description which contains the actual description. Save it all in a new tibble called all_descriptions.

```{r}
all_descriptions <- face_descriptions_clean %>%
  pivot_longer(cols = c("description_1":"description_3"),
               names_to = "t-value",
               values_to = "description") %>%
  select(-'t-value')

all_descriptions
```

** Question 6 **

But, wait… another problem arises… not everyone provided three descriptions! So, if you look at the data, we have some missing values (NAs). We also have some nonsense descriptions (e.g. “f” and “.”). Let’s now sort these out.

Use the filter function to get rid of NA’s and useless descriptions.

Hint Hint: look back at the previous sections where we dealt with null values (i.e. the is.na() function. If you want to keep everything that is not equal to NA, what would you need to do? If you wanted to make sure you kept everything where the description variable had more than one character, is there a function you could use? This is a task extension - you haven’t used this function before, but try googling for a function that counts the number of characters in a variable. You can then use a logical operator (which we also learned about this week), to ensure you only select where there is more than 1 character present.

```{r}
all_descriptions <- all_descriptions %>%
  filter(is.na(description))

all_descriptions

# this question doesn't work, I will wait for the solution
```

