---
title: "R Notebook"
output: html_notebook
---

```{r}
library(CodeClanData)
```

1.1 BEER (dataframe)

Inspect beer data

```{r}
str(beer)
```

View beer data

```{r}
beer
```

Convert calories from chr to num to enable manipulation

```{r}
calories <- as.double(beer$calories)
class(calories)
```

Calculate average percent, calories and carbs from all beers

```{r}
mean(beer$percent)
mean(calories) # the value converted to num, not the chr value in the dataset
mean(beer$carbohydrates)
```

Identify min and max percent beers

```{r}
beer[beer$percent == min(beer$percent) | beer$percent == max(beer$percent), ]
```

Create low/med/high percent subsets

```{r}
low_percent <- subset(beer, percent < 1)
med_percent <- subset(beer, percent >= 1 & percent < 9)
high_percent <- subset(beer, percent > 9)
low_percent
med_percent
high_percent
```

1.2 TEMPERATURES (matrix)

Inspect temp data

```{r}
str(temp)
```

View temp data

```{r}
temp
```

Assign matrix row names

```{r}
rownames(temp) <- c(1910:2015)
temp[1:10, ] # show first 10 rows to show row names
```

Calculate average temperature for each month

```{r}
colMeans(temp)
```

Identify minimum temperature for each month

```{r}
apply(temp, 2, min) # 1 = rows, 2 = columns
```

Identify maximum temperature for each month

```{r}
apply(temp, 2, max)
```

1.3 GAME OF THRONES (nested list)

Inspect game_of_thrones data

```{r}
str(game_of_thrones)
```

Identify number of characters

```{r}
length(game_of_thrones[[10]])
```

Extract all details for character "Walder"

```{r}
game_of_thrones$characters[[1]]
```

