---
title: "R Notebook"
output: html_notebook
---

```{r}
library("dplyr")
```


2 MVP

2.1 Base R cheatsheet

See the file base_r_cheatsheet.rmd

2.2 if statements and functions

2.2.1 Write a function which multiples all numbers by 15 and adds 4

```{r}
multiply_number_by_15_then_add_4 <- function(x){
  print((x * 15) + 4)
}
```

```{r}
multiply_number_by_15_then_add_4(1)
multiply_number_by_15_then_add_4(2)
multiply_number_by_15_then_add_4(3)
```

2.2.2 Write a function that returns “bird”, if the input is “goose”, “duck” or “pigeon”, and returns “not bird” for all other input

```{r}
return_bird_or_not_bird <- function(bird){
  if(bird == "goose" | bird == "duck" | bird == "pigeon"){
    return("bird")
  } else {
    return("not bird")
  }
}
```

```{r}
return_bird_or_not_bird("goose")
return_bird_or_not_bird("apple")
return_bird_or_not_bird(5)
return_bird_or_not_bird("seagull")
```

2.2.3 Write a function which multiplies all numbers by 10. If the input isn’t a number, the function should return “Invalid input”

```{r}
multiply_number_by_10 <- function(x){
  if(is.numeric(x)){
    return(x * 10)
  } else{
    print("Invalid input")
  }
}
```

```{r}
multiply_number_by_10(2)
multiply_number_by_10(5.3)
multiply_number_by_10("x")
multiply_number_by_10(7)
```

2.3 ifelse(), recode() and case_when()

Library dplyr loaded at top of notebook

2.3.1

```{r}
sizes <- c(13, 15, 10, 8, 22)

case_when(
  sizes > 10 ~ "Big",
  sizes <= 10 ~ "Small"
)
```

2.3.2

```{r}
fruit <- c("apple", "apple", "banana", "cherry", "cherry")

recode(fruit, "apple" = "alpaca", "banana" = "bonobo", "cherry" = "chipmunk")
```

2.3.3

```{r}
wife_age <- c(29L, 30L, 36L, 31L, 22L)
husband_age <- c(28L, 35L, 32L, 31L, 29L)

case_when(
  wife_age > husband_age ~ "Wife older",
  husband_age > wife_age ~ "Husband older",
  wife_age == husband_age ~ "Same age"
)
```

2.4 Loops

```{r}
my_list <- list(
  4,
  "apple",
  10.5,
  c("is", "an", "apple"),
  "apple",
  c("i", "like", "apples"),
  8
)
```

2.4.1

```{r}
extract_chars_from_list <- function(item){
  for(item in my_list){
    if(is.character(item)){
      print(item)
    }
  }
}

extract_chars_from_list(my_list)
```

2.4.2

```{r}
extract_chars_from_list_length_greater_than_1 <- function(item){
  for(item in my_list){
    if(is.character(item) & length(item) > 1){
      print(paste(item, collapse = " "))
    }
  }
}

extract_chars_from_list_length_greater_than_1(my_list)
```
